Towards open collaboration in biomedical and health research: An open collaborative scoping review

### 2025-02-11 doc/paper.qmd
docs: :memo: add a draft abstract (not results and conclusion)
Closes #95

---
title: "Towards open collaboration in biomedical and health research: An open collaborative scoping review"
date: last-modified
author:
  - name: Mario Garcia
    affiliations:
      - ref: ku
  - name: Daniel B. Ibsen
    affiliations:
      - ref: sdca
      - ref: au
  - name: Luke W. Johnston
    affiliations:
      - ref: sdca
      - ref: au
affiliations:
  - id: sdca
    name: Steno Diabetes Center Aarhus
    country: Denmark
  - id: au
    name: Aarhus University
    country: Denmark
  - id: ku
    name: Copenhagen University
    country: Denmark
format:
  html: default
  academic-typst:
    section-numbering: "1.1.1"
    columns: 2
    papersize: a4
    margin:
      x: 2cm
      y: 2cm
format-links:
  - html
  - format: academic-typst
    text: PDF
    icon: file-pdf
abstract: |
  {{< lipsum 1 >}}
execute: 
  echo: false
  message: false
  warning: false
---

@@

In Turoman et al the authors describe in detail how the have set up a laboratory whose members collaborate in an open way from study design to dissemination. Here we highlight the role of group meetings as opportunities to propose ideas and have an initial study design proposed. These group meetings are usually started with a section of interesting mistakes, which allows to share the issues, troubleshoot and make the laboratory members comfortable with the fact that errors are unavoidable in science. The feedback from these meetings might lead to updates to a lab manual, which is a great tool for new members to start engaging with the open practices of the lab. This coupled with a support system by senior members of the laboratory allows, in theory, to ease the learning curve of new members to adapt to open practices and collaboration.

In the paper, authors describe their implementation of open practices in each step of the scientific production. Here we focus on the aspects in pre-registration, data analysis and dissemination that improve collaboration between memebers of the labortory and of the same field. For pre-registration they combine slack channels and Trello to communicate between members and OSF for data reposition and uploading pre-registered studies. They put special emphasis on folder structure, which is consistent for all projects, and assigning the authorship in this stage. For data analysis, the authors highlight the figure of the data partner, a member of the laboratory not originally included in the analyses that runs the analyses described in the pre-registered study independently. If the results mismatch with the original ones meetings with the lead analyst or all collaborators are organized to assess what the issue might be. Additionally, this allows for insights that might result in deviations of the original pre-registered study, which should be included and noted in the final manuscript. Importantly, the role of the data partner is described in the final manuscript. Finally, in the dissemination step, they highlight that they might incorporate feedback from presenting their analyses in conferences, which are often included as additional non-preregistered analyses.

Finally, in the paper they also highlight relevant open practices that might not enhance collaboration, but that we consider interesting reviewing too. First, during pre-registration, they allow for papers to contain defined confirmatory and exploratory sections to avoid the usual rigidness of pre-registered studies. Second, in exploratory analyses they included soundness checks whenever possible. Third, the perform sequential hypothesis testing to assess within the limits of a maximum sample size which hypothesis fits the data better without induling in p-hacking. Finally, they collaborate with schools and with universities to disemminate their results.

>>>>

---
title: "Towards open collaboration in biomedical and health research: An open collaborative scoping review"
date: last-modified
author:
  - name: Mario Garcia
    affiliations:
      - ref: ku
  - name: Daniel B. Ibsen
    affiliations:
      - ref: sdca
      - ref: au
  - name: Luke W. Johnston
    affiliations:
      - ref: sdca
      - ref: au
affiliations:
  - id: sdca
    name: Steno Diabetes Center Aarhus
    country: Denmark
  - id: au
    name: Aarhus University
    country: Denmark
  - id: ku
    name: Copenhagen University
    country: Denmark
format:
  html: default
  academic-typst:
    section-numbering: "1.1.1"
    columns: 2
    papersize: a4
    margin:
      x: 2cm
      y: 2cm
format-links:
  - html
  - format: academic-typst
    text: PDF
    icon: file-pdf
abstract: |
  **Background**: Greater, more extensive collaboration is increasingly needed
  in the modern research environment, often to the level of team-based science.
  Combined with this need is the emphasize on doing more open science. However,
  there seems to be less attention on the overlap between open science and
  collaboration.

  **Objectives**: To conduct a scoping review on open collaborative practices used 
  across academia, including methods, tools, and resources.

  **Eligibility criteria**: We included any document that discussed, in any way,
  open collaboration practices, only those in English, and only those that were
  open access. We searched 5 years from whenever we ran a search.

  **Sources of evidence**: We used the OpenAlex database to search for scholarly
  works, which includes a large number of sources. Our search period was
  between 2019-02-28 to 2024-02-29.

  **Charting methods**: All three authors participated in the search, extraction,
  and charting process. As no standard charting form exists for our purposes,
  we developed our own template.

  **Results**: TBD

  **Conclusions**: TBD
execute: 
  echo: false
  message: false
  warning: false
---

@@
